---
title: "STATS_5:Genres+2popularity"
output: html_document
---


```{r import_data}
library(readr)
features <- read_csv("features_distinct.csv")
View(features)
```


```{r remove_duplicates}
library(dplyr)
features1 <- features %>% distinct_at(vars(-ID, -Popularity), .keep_all = TRUE)  # for the match consider all features + year + genre
print("DONE")
```


```{r drop_rows_with_NaN}
features2 <- na.omit(features1)
View(features2)
print("DONE")
```


```{r inspect}
features2$structure_chorus_repetitions
colnames(features2)
```



-----------------------------------------------------------------------------------------------------------------------------------------------------------
-----------------------------------------------------------------------------------------------------------------------------------------------------------
                                                                      ANALYSIS FOR 5 GENRES
-----------------------------------------------------------------------------------------------------------------------------------------------------------
-----------------------------------------------------------------------------------------------------------------------------------------------------------


```{r subset+downsample(acrossGenre):5}
library(ggplot2)
library(dplyr)
library(rstatix)
# select decades
year_selection <- seq(1991, 2020, by=1)
sub_set_all <- features2[which(features2$Date %in% year_selection),]
hist(sub_set_all$Date)
hist(sub_set_all$Popularity)
hist(sub_set_all$lfm_playcount)

# eval genres #
unique(sub_set_all[c("Genre")]) # rap, pop, rock, r-b, country
all_genres <- sub_set_all$Genre
ggplot(data.frame(all_genres), aes(x=all_genres)) + geom_bar()

# select genres #
genres <- c("rap", "pop", "rock", "r-b", "country")
sub_set_all_5genres <- sub_set_all[which(sub_set_all$Genre %in% genres),]

# downsampling #
library(caret)
library(groupdata2)
down_sub_set_all5 <- sub_set_all_5genres
down_sub_set_all5$Date2 <- paste(down_sub_set_all5$Date, down_sub_set_all5$Genre, sep="") # add ID column
set.seed(27)
down_sub_set_all5 <- downsample(down_sub_set_all5, cat_col = "Date2")
hist(down_sub_set_all5$Date)
hist(down_sub_set_all5$Popularity)
hist(down_sub_set_all5$lfm_playcount)
all_genres <- down_sub_set_all5$Genre
ggplot(data.frame(all_genres), aes(x=all_genres)) + geom_bar()
```


```{r count}
samples_genre <- which(all_genres == "pop")
length(samples_genre)
```


```{r get_subSet}
write.csv(down_sub_set_all5, "5Genres_SubSet.csv", row.names = TRUE)
```


```{r multiple_regression_date_(initial_model)}
library(lm.beta)
library(olsrr)
library(car)
library(ggplot2)
library(usdm)
library(dplyr)

# Fit model through lm(outcome ~ predictor)
sub_set5 <- subset(down_sub_set_all5, select=-c(ID, Date2)) # drop not used variables
sub_set5[,which(!names(sub_set5) %in% c("Date", "Genre"))] <- scale(sub_set5[,which(!names(sub_set5) %in% c("Date", "Genre"))]) # Z-score normalisation for the IV
sub_set5A <- sub_set5

for (var in colnames(sub_set5)){
  if (sum(is.na(sub_set5[var])) > 0){
    print(var)
    drops <- c(var)
    sub_set5A <- sub_set5A[ , !(names(sub_set5A) %in% drops)]
  }
}

# drop all emotion-related features 
todrop <- c("AFINN", "BingLiu", "MPQA", "NRCAffect", "NRCEmotion", "NRCExpanded", "NRCHashtag", "Sentiment140", "SentiWordNet", "LabMT", "Negation_count", "six_letter_words", "LIWC")
sub_set5Ab <- sub_set5A[, -grep(paste(todrop, collapse="|"), colnames(sub_set5A))]
# Add emotion-related features from "LIWC"
final_features <- c(colnames(sub_set5Ab), c("LIWC_anger", "LIWC_anx", "LIWC_sad", "LIWC_negemo", "LIWC_posemo"))
sub_set5B <- sub_set5A[,c(final_features)]

lm.model5 <- lm(Date ~. -Genre, data=sub_set5B)
summary(lm.model5)
```


```{r fix_Collinearity}
# We consider features with a variance inflation factor (VIF) below 5 to prevent multicollinearity
sub_set5B <- subset(sub_set5B, select=-c(lexical_punctuation_count, lexical_hapax_legomenon_ratio, structure_number_of_verses, Genre)) # Remove variables that return NA + categorical

vif.1 <- usdm::vif(as.data.frame(sub_set5B))
vif.1

for (y in seq(1:length(vif.1[,1]))){
  max.to.drop <- vif.1[which.max(vif.1[,2]),]
  if (max.to.drop[,2] > 5){
    sub_set5B <- dplyr::select(sub_set5B, -max.to.drop[,1])
    vif.1 <- usdm::vif(sub_set5B)
    print("Iterating ...")
  }
}

vif.2 <- usdm::vif(sub_set5B)
vif.2
new_sub_set5 <- sub_set5B
```



```{r fix_Kooc's_distance}
##  DROP FEATURES WITH COOK'S DISTANCE > 0.5  ##

sub_set_all5 <- new_sub_set5
new_sub_set5B <- sub_set_all5

for (feature in colnames(sub_set_all5)) {
  #if (feature != "Date"){
  if (feature != "Date" & feature != "Popularity"){
    each.model <- lm(Date ~ eval(parse(text = feature)), data=sub_set_all5)
    cooksd <- cooks.distance(each.model)
    if (max(cooksd) > 0.5){
      print(feature)
      new_sub_set5B <- subset(new_sub_set5B, select=-c(eval(parse(text = feature))))
    }
  }
}
print("DONE")
```


```{r get_final_data}
library(nlme)
library(ggplot2)
# Get filtered data.frame with genre information
sub_set5$row_index <- 1:nrow(sub_set5)
sub_set_all5$row_index <- 1:nrow(sub_set_all5)
genr_col <- sub_set5[which(sub_set5$row_index %in% sub_set_all5$row_index),]
sel_Data5 <- cbind(new_sub_set5B, Genre=genr_col$Genre)
# run initial model
lm.model5.0 <- lm(Date ~ . -Genre, data=sel_Data5)
plot(lm.model5.0)
```



In order to investigate whether there is a a relationship between Popularity and Genre, as well as the interaction between Popularity:Date and Genre,
we start by performing multiple logistic regression considering genre as Dependent Variable and Popularity and Popularity:Date as predictors.

```{r multiple_logistic_regression}
library(mlogit)
library(car)
library(effects)
library(multcomp)
#Final_Data5 <- subset(sel_Data5, select=-c(row_index)) 
Final_Data5 <- sel_Data5
is.factor(Final_Data5$Genre)

Final_Data5B <- mlogit.data(Final_Data5, choice ="Genre", shape = "wide")
mlogit.model5 <- mlogit(Genre ~ 1 | Date*Popularity, data=Final_Data5B, reflevel = "rap")
summary(mlogit.model5)
exp(mlogit.model5$coefficients)
exp(confint(mlogit.model5))
```


We perform the same analysis considering playcount instead of popularity.

```{r multiple_logistic_regression_playcount}
mlogit.model5.playcount <- mlogit(Genre ~ 1 | Date*lfm_playcount, data=Final_Data5B, reflevel = "rap")
summary(mlogit.model5.playcount)
exp(mlogit.model5.playcount$coefficients)
exp(confint(mlogit.model5.playcount))
```


To investigate the relevant features to predict changes in lyrics over time, multiple linear regression (for each genre individually) is carried out considering year as dependent variable. Since popularity showed to vary across genres, both the lyrics features individually as well as their interaction with popularity will be investigated. 
Each model was optimised following backwards selection.


```{r descriptive}
library(psych)
describeBy(Final_Data5, Final_Data5$Genre)
```


```{r linear_models_standard}
# select genre
sub_set_rap5 <- Final_Data5[which(Final_Data5$Genre == 'rap'),]
subset_rap5 <- subset(sub_set_rap5, select=-c(Genre))
sub_set_rock5 <- Final_Data5[which(Final_Data5$Genre == 'rock'),]
subset_rock5 <- subset(sub_set_rock5, select=-c(Genre))
sub_set_pop5 <- Final_Data5[which(Final_Data5$Genre == 'pop'),]
subset_pop5 <- subset(sub_set_pop5, select=-c(Genre))
sub_set_rb5 <- Final_Data5[which(Final_Data5$Genre == 'r-b'),]
subset_rb5 <- subset(sub_set_rb5, select=-c(Genre))
sub_set_country5 <- Final_Data5[which(Final_Data5$Genre == 'country'),]
subset_country5 <- subset(sub_set_country5, select=-c(Genre))

lm.rap5 <- lm(Date ~ ., data=subset_rap5)
lm.rock5 <- lm(Date ~ ., data=subset_rock5)
lm.pop5 <- lm(Date ~ ., data=subset_pop5)
lm.rb5 <- lm(Date ~ ., data=subset_rb5)
lm.country5 <- lm(Date ~ ., data=subset_country5)
print("DONE")
```

```{r explore_features}
colnames(Final_Data5)
```


```{r linear_models_withinGenre}
library(stringr)
sel.model.rap5 <- step(lm.rap5, direction = "backward", test = "F")
toselect.x.rap5 <- summary(sel.model.rap5)$coeff[-1,4] < 1 # Get all predictors (we do not filter by significant level)
relevant.x.rap5 <- names(toselect.x.rap5)[toselect.x.rap5 == TRUE]
interaction.formula.rap5 <- as.formula(paste("Date ~", paste(relevant.x.rap5, collapse= "+Popularity*")))
interaction.model.rap5 <- lm(formula=interaction.formula.rap5, data=subset_rap5)
interaction_lfm.formula.rap5 <- as.formula(paste("Date ~", paste(relevant.x.rap5, collapse= "+lfm_playcount*")))
interaction_lfm.model.rap5 <- lm(formula=interaction_lfm.formula.rap5, data=subset_rap5)

sel.model.rock5 <- step(lm.rock5, direction = "backward", test = "F")
toselect.x.rock5 <- summary(sel.model.rock5)$coeff[-1,4] < 1 
relevant.x.rock5 <- names(toselect.x.rock5)[toselect.x.rock5 == TRUE] 
interaction.formula.rock5 <- as.formula(paste("Date ~", paste(relevant.x.rock5, collapse= "+Popularity*")))
interaction.model.rock5 <- lm(formula=interaction.formula.rock5, data=subset_rock5)
interaction_lfm.formula.rock5 <- as.formula(paste("Date ~", paste(relevant.x.rock5, collapse= "+lfm_playcount*")))
interaction_lfm.model.rock5 <- lm(formula=interaction_lfm.formula.rock5, data=subset_rock5)

sel.model.pop5 <- step(lm.pop5, direction = "backward", test = "F")
toselect.x.pop5 <- summary(sel.model.pop5)$coeff[-1,4] < 1
relevant.x.pop5 <- names(toselect.x.pop5)[toselect.x.pop5 == TRUE] 
interaction.formula.pop5 <- as.formula(paste("Date ~", paste(relevant.x.pop5, collapse= "+Popularity*")))
interaction.model.pop5 <- lm(formula=interaction.formula.pop5, data=subset_pop5)
interaction_lfm.formula.pop5 <- as.formula(paste("Date ~", paste(relevant.x.pop5, collapse= "+lfm_playcount*")))
interaction_lfm.model.pop5 <- lm(formula=interaction_lfm.formula.pop5, data=subset_pop5)

sel.model.rb5 <- step(lm.rb5, direction = "backward", test = "F")
toselect.x.rb5 <- summary(sel.model.rb5)$coeff[-1,4] < 1
relevant.x.rb5 <- names(toselect.x.rb5)[toselect.x.rb5 == TRUE] 
interaction.formula.rb5 <- as.formula(paste("Date ~", paste(relevant.x.rb5, collapse= "+Popularity*")))
interaction.model.rb5 <- lm(formula=interaction.formula.rb5, data=subset_rb5)
interaction_lfm.formula.rb5 <- as.formula(paste("Date ~", paste(relevant.x.rb5, collapse= "+lfm_playcount*")))
interaction_lfm.model.rb5 <- lm(formula=interaction_lfm.formula.rb5, data=subset_rb5)

sel.model.country5 <- step(lm.country5, direction = "backward", test = "F")
toselect.x.country5 <- summary(sel.model.country5)$coeff[-1,4] < 1
relevant.x.country5 <- names(toselect.x.country5)[toselect.x.country5 == TRUE] 
interaction.formula.country5 <- as.formula(paste("Date ~", paste(relevant.x.country5, collapse= "+Popularity*")))
interaction.model.country5 <- lm(formula=interaction.formula.country5, data=subset_country5)
interaction_lfm.formula.country5 <- as.formula(paste("Date ~", paste(relevant.x.country5, collapse= "+lfm_playcount*")))
interaction_lfm.model.country5 <- lm(formula=interaction_lfm.formula.country5, data=subset_country5)

print("DONE")
```


We do not consider the interaction with popularity for any genre as it does not improve the model significantly (p<.01).

```{r final_models}
# 3-factor anova (compariosn between: baseline, selected, and interaction models)
anova(sel.model.rap5, interaction.model.rap5)
anova(sel.model.rap5, interaction_lfm.model.rap5)
anova(sel.model.rock5, interaction.model.rock5)
anova(sel.model.rock5, interaction_lfm.model.rock5)
anova(sel.model.pop5, interaction.model.pop5)
anova(sel.model.pop5, interaction_lfm.model.pop5)
anova(sel.model.rb5, interaction.model.rb5)
anova(sel.model.rb5, interaction_lfm.model.rb5)
anova(sel.model.country5, interaction.model.country5)
anova(sel.model.country5, interaction_lfm.model.country5)

# We select the best model for each genre
plot(sel.model.rap5)
plot(sel.model.rock5)
plot(sel.model.pop5)
plot(sel.model.rb5)
plot(sel.model.country5)

summary(sel.model.rap5)
summary(sel.model.pop5)
summary(sel.model.rock5)
summary(sel.model.rb5)
summary(sel.model.country5)
```


```{r check_relationships}
library(ggplot2)
# have  look to ALL realtionships between DV & IV without transformation
# RAP
for (feature in relevant.x.rap5) {
  print(ggplot(sub_set_rap5, aes(y=eval(parse(text=feature)), x=Date)) +
  geom_smooth(method = "lm", aes(group = 1)) + geom_jitter(alpha = .5) + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ggtitle(feature))
}

#ROCK
for (feature in relevant.x.rock5) {
  print(ggplot(sub_set_rock5, aes(y=eval(parse(text=feature)), x=Date)) +
  geom_smooth(method = "lm", aes(group = 1)) + geom_jitter(alpha = .5) + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ggtitle(feature))
}

#POP
for (feature in relevant.x.pop5) {
  print(ggplot(sub_set_pop5, aes(y=eval(parse(text=feature)), x=Date)) +
  geom_smooth(method = "lm", aes(group = 1)) + geom_jitter(alpha = .5) + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ggtitle(feature))
}

#RB
for (feature in relevant.x.rb5) {
  print(ggplot(sub_set_rb5, aes(y=eval(parse(text=feature)), x=Date)) +
  geom_smooth(method = "lm", aes(group = 1)) + geom_jitter(alpha = .5) + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ggtitle(feature))
}

#COUNRTY
for (feature in relevant.x.country5) {
  print(ggplot(sub_set_country5, aes(y=eval(parse(text=feature)), x=Date)) +
  geom_smooth(method = "lm", aes(group = 1)) + geom_jitter(alpha = .5) + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ggtitle(feature))
}
```


-----------------------------------------------------------------------------------------------------------------------------------------------------------

-----------------------------------------------------------------------------------------------------------------------------------------------------------
